Do More Subdivisions in Wards Increase Voter Turnout? An Analysis of the 2022 Toronto Municipal Election

### 2023-06-03 outputs/paper/toronto_elections_paper.qmd
Fixing typos and changing phrasing.

#Voter turnout by ward
#Code refernced from: https://stackoverflow.com/questions/11335836/increase-number-of-axis-ticks
finalized_voter_statistics |>
 ggplot(mapping = aes(x = ward, y = turnout)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 25)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 25))+
  geom_bar(stat = "identity") +
  labs(
    title = "Voter Turnout by Ward",
    x = "Ward",
    y = "Turnout (%)",
  ) +
  theme_classic() 
```
### Comparing Number of Subdivisions and Turnout

@@

There are several limitations with conducting this analysis. The creation of subdivision boundaries within wards lacks transparency and clarity, while impacting the local democratic process. The City Clerk is responsible for drawing up all ward subdivisions, however the data and information they use to make these decisions is not public and it is challenging to account for any bias introduced during this process. Toronto Elections @ElxnAccessibility clearly outlines their process for finding and establishing accessible polling locations within subdivisions but does not elaborate on the process of establishing subdivisions themselves.

@@

## Conclusion
This paper investigates the number of subdivisions within Toronto’s 25 wards in relation to voter turnout to see if the wards with more subdivisions had higher turnout during the 2022 municipal election. My results reveal that wards with more subdivisions did not correlate to having higher voter turnout. Factors such as income, level of education attained, and an analysis of local races for councillor and school board trustee within a ward (whether there was an incumbent or not) should be analyzed in further studies of voter behaviour during municipal elections. Data from the upcoming Toronto Mayoral by-election on June 26, 2023 may reveal further insights into the placement of polling locations in relation to voter turnout. Cities and municipalities across Ontario are facing declining voter turnout and the factors must be identified and addressed, whether it be more strategically choosing subdivisions and polling locations in wards, introducing online voting mechanisms, or implementing democratic reform [@reducingcost]. Strong local representation is crucial for the future of democracy in Toronto and subdivisions play an important role in ensuring voters have their voices heard.

>>>>

#Voter turnout by ward
#Code refernced from: https://stackoverflow.com/questions/11335836/increase-number-of-axis-ticks
finalized_voter_statistics |>
 ggplot(mapping = aes(x = ward, y = turnout)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 25)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 25))+
  geom_bar(stat = "identity") +
  labs(
    title = "Voter Turnout by Ward",
    x = "Ward",
    y = "Turnout (%)",
  ) +
  theme_classic() 
```

@@

There are several limitations with conducting this analysis. The creation of subdivision boundaries within wards lacks transparency and clarity, while impacting the local democratic process. The City Clerk is responsible for drawing up all subdivisions, however the data and information they use to make these decisions is not public and it is challenging to account for any bias introduced during this process. Toronto Elections @ElxnAccessibility clearly outlines their process for finding and establishing accessible polling locations within subdivisions but does not elaborate on the process of establishing subdivisions themselves.

@@

## Conclusion
This paper investigates the number of subdivisions within Toronto’s 25 wards in relation to voter turnout to see if the wards with more subdivisions had higher turnout during the 2022 municipal election. My results reveal that wards with more subdivisions did not correlate to having higher voter turnout. Factors such as income, level of education attained, and an analysis of local races for councillor and school board trustee within a ward should be analyzed in further studies of voter behaviour during municipal elections. Data from the upcoming Toronto Mayoral by-election on June 26, 2023 may reveal further insights into the placement of polling locations in relation to voter turnout. Cities and municipalities across Ontario are facing declining voter turnout and the factors must be identified and addressed, whether it be more strategically choosing subdivisions and polling locations in wards, introducing online voting mechanisms, or implementing democratic reform [@reducingcost]. Strong local representation is crucial for the future of democracy in Toronto and subdivisions play an important role in ensuring voters have their voices heard.
